\documentclass[aspectratio=169]{beamer}
\usepackage{standalone}

\usepackage{stmaryrd}
\usepackage{listings}
\usepackage{bussproofs}

\usepackage[hyperref=auto,style=alphabetic,backend=bibtex]{biblatex}
\addbibresource{kwarcpubs.bib}
\addbibresource{extpubs.bib}
\addbibresource{extcrossrefs.bib}
\usepackage{appendixnumberbeamer}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{mmt}
\usetikzlibrary{docicon}

\usetheme{Pittsburgh}
\setbeamertemplate{footline}[frame number]
\setbeamertemplate{navigation symbols}{}
\usecolortheme{beaver}
\setbeamertemplate{frametitle}[default][left]
% \setbeamersize{text margin left=3em}

\usepackage{utils/colors}
\usepackage[forbeamer]{utils/basic}
\usepackage{utils/operators}
\usepackage{utils/mylstmisc}
\usepackage{utils/lstmmt}

\lstset{basicstyle=\ttfamily}
\lstset{commentstyle=\itshape\color{commentfont}}

\title{\textit{What does it mean?}\\A framework for prototyping Montague-style semantics}

\author{Jan Frederik Schaefer}
\institute{FAU Erlangen-N\"urnberg/KWARC}
\date{\textbf{Seminar: Computing Meaning}\\Hildesheim\\July 21, 2022}

\begin{document}
\begin{frame}[fragile, noframenumbering]
    \textbf{What does it mean? A framework for prototyping Montague-style semantics}
    \small

    \begin{block}{}
        \parbox{\linewidth}{
            One approach to study the meaning of natural language is to translate
            utterances to logical expressions, which have well-defined semantics and allow
            for rigorous inference.  Richard Montague, a well-known pioneer of this
            approach, described the translation of a particular subset of English into
            logic in great detail.  Over the years, there have been many similar
            publications that propose a novel way to translate sentences into a logic
            (often first-order logic, but more obscure logics are also common).  Sometimes,
            the approach was extended with pragmatic analysis based on logical inference.

            We claim that research (and education) in this direction could benefit from
            more prototyping to test and demonstrate new ideas.

            In this talk, I will present GLIF, a declarative framework for prototyping the
            translation of natural language to logics.  GLIF combines existing, specialized
            frameworks that solve part of the problem: the Grammatical Framework
            (development of natural language grammars), MMT (logic development) and ELPI
            (inference).  These frameworks can be connected seamlessly because of their
            compatible underlying logical frameworks.  We successfully use GLIF in a
            lecture on symbolic natural language semantics.
        }
    \end{block}
\end{frame}

\frame\titlepage

\begin{frame}
    \frametitle{whoami}
    \begin{itemize}
        \item PhD student in the KWARC group (Erlangen)
            \com{Knowledge representation}
        \item Supervisor: Michael Kohlhase
        \item My interest: Mathematical language, precise semantics extraction
        \item Background for this talk:
            \begin{itemize}
                \item We teach a lecture in \emph{logic-based natural language semantics}
                \item Wanted more hands-on experience
                \item $\leadsto$ new framework: GLIF\com{Grammar, Logic, Inference}
            \end{itemize}
    \end{itemize}
\end{frame}

% \input{slides/deep-nlu}


\input{slides/fragments-to-glif-2}

\setlength\pdfpagewidth{10cm}%
\setlength\pdfpageheight{9cm}%
\setlength\textwidth{8cm}
% \newgeometry{left=0.1cm,bottom=0.1cm}

{
    \enablepart{showjupyter}
    \disablepart{showepistemicexample}
    \disablepart{mentionprovergen}
    \input{slides/glif-components}
}

% \input{slides/glif-example-tab-machine}
% 
% \input{slides/glif-example-translation}
% 
% 
% \begin{frame}
%     \frametitle{GLIF Summary}
%     \includestandalone[width=\textwidth]{fig/glif-architecture}
% \end{frame}
% 
% \begin{frame}
%     \frametitle{Supporting the Semantic/Pragmatic Analysis}
%     Observation:
%     \begin{itemize}
%         \item Parsing and semantics construction are based on specialized frameworks
%         \item ELPI is a more general programming language
%     \end{itemize}
%     
%     \vspace{1em}
%     Can we do something more specialized?
%     \begin{itemize}
%         \item Not really -- there is no ``standard recipe'' for
%             semantic/pragmatic analysis
%         \item But: It usually involves inferential reasoning
%     \end{itemize}
% 
%     \vspace{2em}
%     \centering
%     \bfseries Let's generate provers!\par
% \end{frame}
% 
% \input{slides/elpi-gen}
% 
% \begin{frame}
%     \frametitle{Conclusion: Prototyping NLU Pipelines}
%     \includestandalone[width=\textwidth]{fig/glif-spec}
% \end{frame}

\appendix

\begin{frame}[allowframebreaks,t]
    \frametitle{References}
    \printbibliography
\end{frame}

\end{document}
